Characterizing Water Composition with an Autonomous Robotic Team Employing Comprehensive In-Situ Sensing, Hyperspectral Imaging, Machine Learning, and Conformal Prediction

### 2024-02-23 paper/paper.tex
updated text

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{
Inland waters pose a unique challenge for water quality monitoring by remote sensing techniques due to their complicated spectral features and small-scale variability. At the same time, collecting the high-quality reference data needed to calibrate remote sensing data products is both time consuming and expensive. In this study, we present the further development of a robotic team composed of an uncrewed surface vessel (USV) providing in situ reference measurements and an unmanned aerial vehicle (UAV) equipped with a hyperspectral imager. Together, this team is able to address the limitations of existing approaches by enabling the simultaneous collection of comprehensive in situ data together with high-resolution hyperspectral imagery. We showcase the capabilities of this team using data collected in a north Texas pond during three collections in the fall of 2020. Machine learning models for 13 variables are trained using the data set of paired in situ measurements and coincident reflectance spectra. These models can successfully estimate physical variables, including temperature, conductivity, pH, and turbidity, as well as the concentrations of blue-green algae, colored dissolved organic matter (CDOM), chlorophyll-a, crude oil, optical brighteners and the ions $\mathrm{Ca}^{2+}$, $\mathrm{Cl}^{-}$, and $\mathrm{Na}^{+}$. We extend the training procedure to utilize conformal prediction to estimate 90\% confidence intervals for the output of each trained model. The maps generated by applying each trained model to the collected images reveal the small-scale spatial variability within the pond. Additionally, the permutation importance computed for each feature of the trained models highlights the spectral features relevant to each water quality variable.
}

@@

For decades, remote sensing imagery has been used for environmental monitoring with applications ranging from resource mapping, land type classification, and urban growth assessment to wildfire tracking, natural disaster tracking, and many more \cite{cohen2004landsat, melesse2007remote, allison2016airborne, joyce2009review, robert2017analysis}. Among these applications, the retrieval of water quality variables from remote sensing imagery remains challenging due to the difficulty of obtaining in situ reference data that coincide with available satellite imagery. Many studies have shown success modeling relevant water quality variables such as colored dissolved organic matter (CDOM), cholorophyll-a, and total suspended sediment concentrations by using combinations of spectral bands from terrestrial remote sensing satellites. However, the in situ reference data used in these studies typically require the time-consuming collection of individual samples for lab analysis or continuous monitoring at fixed sites \cite{brezonik2005landsat, paavel2011optical, gursoy2019investigating, bonansea2019using, sagan2020monitoring, remote-sensing-finland}.

Much effort has recently been spent on the curation of comprehensive datasets that combine water quality records with decades of satellite imagery to enable the development of new methods for monitoring water quality. For example, Aurin et al. curated over 30 years of oceanographic field campaign data with associated coincident satellite imagery \cite{aurin2018remote}. Similarly, Ross et al. have combined more than 600,000 records of dissolved organic carbon, chlorophyll-a and other variables of water quality together with Landsat historical reflectance data for the period 1984--2019 \cite{ross2019aquasat}. With a sufficient quantity of data, machine learning methods can augment traditional approaches using reflectance band ratios to enable estimation of water quality metrics and perform water quality classification \cite{schiller1999neural, brown2008neural, lary2010artificial, feng2015urban,belgiu2016random}. However, the low spatial and spectral resolution of available multiband remote sensing satellites makes it difficult to analyze inland waters with small spatial features and complicated spectral signatures. Furthermore, the long return times of these satellites limit their use for real-time monitoring, for example, during oil spills \cite{fingas2017review}.

@@

Despite the increased spectral and spatial resolution afforded by UAV platforms, these improvements alone cannot address the limited spatial resolution of in situ reference data.  To address this gap and enable comprehensive real-time water quality detection, we have developed a robotic team composed of an autonomous uncrewed surface vessel (USV) equipped with a collection of reference grade instruments together with a UAV carrying a hyperspectral imager \cite{robotTeam1}. By incorporating these reference instruments on a maneuverable USV platform, we are able to rapidly collect large volumes of water quality data for a variety of physical, biochemical, and chemical variables. Critically, the USV enables the collection of reference data with significantly improved spatial resolution compared to other approaches. These in situ measurements then provide the ground-truth data used to train machine learning models which map reflectance spectra captured by the hyperspectral imager to the desired water quality variables.

In this study, we expand on our previous work by demonstrating the capabilities of this coordinated robot team using data collected at a pond in North Texas over three separate days in the fall of 2020. Specifically, we show the ability to accurately predict physical variables such as temperature, conductivity, pH, and turbidity, as well as biochemical constituents, including blue-green algae pigments, CDOM, and chlorophyll-a, in addition to concentrations of crude oils, optical brighteners, and a variety of ions. In addition, the considerable size of the collected data set allows us to quantify the uncertainty in the resulting machine learning models by using conformal prediction techniques. This methodology addresses the shortcomings of existing approaches by facilitating rapid, real-time data collection coupled with on-board hyperspectral image processing and machine learning model deployment with uncertainty quantification. The resulting models are small enough to be deployed on board the UAV to enable real-time streaming of data products to the ground. This approach drastically reduces the time from data collection to insight.

@@

\begin{figure}
\includegraphics[width=\columnwidth]{paper/figures/materials-and-methods/annotated-boat.pdf}
\caption{ Configuration of the USV: (\textbf{a}) Frontal view of the USV showing the Eureka Manta+40 Multiprobes mounted on the underside of the boat. (\textbf{b}) The USV deployed in the water.\label{fig:boat-components}}
\end{figure}

@@

\begin{table}[H] 
\caption{In-situ reference sensors utilized in this study.\label{tab:sensors}}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{CCCCC}
\toprule
\textbf{Sensor}	& \textbf{Units} & Resolution & \textbf{Sensor Type} & \textbf{Target Category}\\
\midrule
Temperature		                    & $^{\circ}$C        & 0.01  & Thermistor                           & Physical \\
Conductivity                        & $\mu$S/cm          & 0.01  & Four-Electrode Graphite Sensor       & Physical  \\
pH                                  & logarithmic (0-14) & 0.01  & Flowing-junction Reference Electrode & Physical \\
Turbidity                           & FNU                & 0.01  & Ion-Selective Electrode              & Physical \\
$\mathrm{Ca^{2+}}$                  & mg/l               & 0.1   & Ion-Selective Electrode              & Ions \\
$\mathrm{Cl^-}$                     & mg/l               & 0.1   & Ion-Selective Electrode              & Ions \\
$\mathrm{Na^+}$                     & mg/l               & 0.1   & Ion-Selective Electrode              & Ions \\
Blue-Green Algae (phycoerythrin)    & ppb                & 0.01  & Fluorometer                          & Biochemical \\
Blue-Green Algae (phycocyanin)      & ppb                & 0.01  & Fluorometer                          & Biochemical \\
CDOM                                & ppb                & 0.01  & Fluorometer                          & Biochemical \\
Chlorophyll A                       & ppb                & 0.01  & Fluorometer                          & Biochemical \\
Optical Brighteners                 & ppb                & 0.01  & Fluorometer                          & Chemical \\
Crude Oil                           & ppb                & 0.01  & Fluorometer                          & Chemical\\
\bottomrule
\end{tabularx}
% \noindent{\footnotesize{\textsuperscript{1} Tables may have a footer.}}
\end{table}

As shown in Table~\ref{tab:sensors}, the physical measurements and ion sensors are largely based on different electrode configurations, while the chemical and biochemical measurements are all optically significant in UV and visible light, enabling their determination by fluorometry. The pigments phycoerythrin and phycocyanin are used to determine the blue-green algae content, which together with chlorophyll-a enable us to assess the distribution of photosynthetic life in the pond. In addition, we also measured the concentration of colored dissolved organic matter (CDOM), which impacts light penetration and serves as a primary source of bioavailable carbon. Crude oil (natural petroleum) and optical brightener concentrations are also measured with fluorometers and are relevant for identifying sources of industrial contamination or natural seepage.

@@

\begin{figure}[H]
\includegraphics[width=\columnwidth]{paper/figures/materials-and-methods/annotated-drone.pdf}
\caption{Configuration of the UAV: (\textbf{a}) the hyperspectral imager and acquisition computer. (\textbf{b}) the assembled UAV with secondary processing computer and (upward facing) downwelling irradiance spectrometer. \label{fig:drone-components}}
\end{figure}

@@

\begin{figure}[H]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=14cm]{paper/figures/materials-and-methods/pipeline-figure-2.pdf}
\end{adjustwidth}
\caption{Hyperspectral Image Processing: Hyperspectral data cubes are collected one scan-line at a time (left). By utilizing downwelling irradiance spectra we convert each pixel from spectral radiance to reflectance. By using orientation and position data from the on-board GPS and INS, we georeference each pixel to assign it a latitude and longitude on the ground. The final data product is a georectified hyperspectral reflectance data cube visualized at the right. \label{fig:hsi-pipeline}}
\end{figure}

@@

\begin{figure}[H]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=15.5cm]{paper/figures/materials-and-methods/HyperSpectralInfoGraphic.pdf}
\end{adjustwidth}
\caption{A georectified reflectance data cube is visualized (center) with the $\log10$-reflectance along the z axis and a pseudo-color image on the top. In the top left we visualize the downwelling irradiance spectrum (the incident light). The surrounding plots showcase exemplar pixel reflectance spectra for open water, dry grass, algae, and a rhodamine dye plume used to test the system.\label{fig:hsi-infographic}}
\end{figure}

@@

Data were collected using the robot team at a site in North Texas during three observation periods in the fall of 2020. For each collection period, the UAV first completed a broad survey of the pond, capturing multiple hyperspectral data cubes. Subsequently, the USV sampled across the pond, collecting in situ reference measurements. Each of these reference measurements was then collocated with individual pixel spectra, where the USV track overlapped with UAV pixels. To account for any time lag in the values measured by the in situ instruments and to account for the USV size in comparison to the data cube's spatial resolution, each in-situ measurement is associated with a 3$\times$3 grid of HSI pixels, that is, a 30 cm $\times$ 30 cm square. These combined data form the tabular data set on which we train regression models; pixel spectra form input features, and each separate USV sensor forms a distinct target variable.

Each collection was performed near solar noon to maximize the amount of incident sunlight illuminating the water; however, to account for any potential variation in lighting conditions between data cubes, we augment the training set with additional features including the drone viewing geometry (roll, pitch, heading) and solar illumination geometry (solar azimuth, solar elevation, solar zenith) as well as the total at-pixel intensity before reflectance conversion, the total downwelling intensity, and the drone altitude. Further feature engineering is performed to add additional spectral indices that utilize combinations of specific wavelength bands such as the normalized difference vegetation index (NDVI), normalized difference water index (NDWI), simple ratio (SR), photochemical reflectance index (PRI) and more, as outlined in \cite{envi_vegetation_indices, thenkabail2018hyperspectral,kaufman1992atmospherically, SpectralIndexWheat}. A comprehensive list of these added features is provided in Supplementary Table S1. The final dataset includes a total of 526 features (462 reflectance bands plus 64 additional features) with over 120,000 records.

@@

For each of the 13 target variables listed in Table~\ref{tab:sensors}, the data was randomly partitioned into a 90:10 training/testing split. Using Julia's MLJ (Machine Learning Julia) package, a variety of machine learning models compatible with the regression task were identified \cite{MLJ1}. This selection of models included linear regression (and it's variations), Gaussian process regression, shallow neural networks, and decision tree-based models. Out of this set, the Random Forest Regressor (RFR) model from \cite{decision-trees} was selected for use in this study due to its relatively quick training and inference times. Random forests are an ensembling technique based on bagged predictions of individual decision tree regressors trained using the classification and regression trees (CART) algorithm. Each individual tree in a RFR is trained on a random subset of features \textit{and} a random subset of training records \cite{decision-trees, random-forest}.

As reflectance values in adjacent wavelength bins tend to correlate with each other and, therefore, may not necessarily contribute additional information content to the final model, it is desirable to evaluate the relative importance of each feature to the final predictions of our trained models. This is useful both for identifying the most relevant features and for performing feature selection to reduce the final model size and optimize the inference time. By default, tree-based methods, such as the RFR allow for an impurity-based ranking as described in \cite{random-forest, rfr-importance-ranking}. However, these methods have been shown to be biased towards high cardinality and correlated features \cite{strobl2008conditional}. Therefore, we choose to use the model-agnostic permutation importance as described by \cite{parr2018beware}. To do this, we further partition the training data set, resulting in an 80:10:10 split with 80\% of points used for model training, 10\% of points used for validation and determination of feature importance, and the final 10\% held out as an independent test data set. The importance of the $j^{\text{th}}$ feature is then computed as
\begin{equation}
    \text{Imp}_j = R^2(f(X_{\text{val}}), y_{\text{val}}) - R^2(f(X_{\text{val}}^{(j)}), y_{\text{val}})
\end{equation}
where $(X_{\text{val}}, y_{\text{val}})$ is the validation set, $f(\cdot)$ is the trained model, $R^2(\cdot, \cdot)$ is the coefficient of determination, and $X_{\text{val}}^{(j)}$ is validation feature set with the $j^{th}$ column randomly permuted. The permutation importance is therefore understood as the decrease in model performance when the $j^{th}$ feature is replaced by random values from the validation set.

To assess the uncertainty of the final model's predictions, we employ Inductive Conformal Prediction as described in \cite{conformal-prediction-1, conformal-prediction-2, conformal-prediction-3, conformal-prediction-4}. To do this, we compute a set of nonconformity scores of the trained model on the validation set using an uncertainty heuristic, in this case, the absolute error
\begin{equation}
    s_i = s(X_i, y_i) = \lvert f(X_i) - y_i \rvert = \lvert \hat{y}_i - y_i \rvert
\end{equation}
where $f(X_i)=\hat{y}_i$ denotes the trained ML model applied to the $i^{th}$ calibration record. These n-many scores are sorted, and the half width of the interval $d$ is calculated as the $\frac{\lceil(1-\alpha)(n+1) \rceil}{n}$ quantile of this set in order to achieve a coverage of $1-\alpha$ on the calibration set. Prediction intervals for new data are then formed as $f(X)\pm d$. For this study, we have chosen $\alpha=0.1$ for a coverage corresponding to a 90\% confidence interval.

Using these tools, the training procedure for each model is as follows: First, the model is trained using six-fold cross-validation on the training set with default hyperparameter values. Feature importances for the trained model are then computed, and the top 25 features are identified. A second model is then trained using the same 6-fold cross-validation scheme with \textit{ only} these 25 most important features together with hyperparameter optimization using a random parameter search over the number of trees and sampling fraction. Model performance is evaluated by computing out-of-fold values for the coefficient of determination, as well as the
\begin{align}
    \text{RMSE} &= \sqrt{\frac{\sum\limits_{i=1}^N (\hat{y}_i-y_i)^2}{N}}, \\
    \text{MAE} &= \frac{\sum\limits_{i=1}^N \lvert \hat{y}_i - y_i \vert}{N},
\end{align}
where RMSE is the root-mean-square error, and MAE is the mean absolute error between true values $y_i$ and predictions $\hat{y}_i$.

Having identified the model with the best out-of-fold performance, we proceed to train the final hyperparameter optimized model on the full training set with associated uncertainty estimated using conformal prediction.  Then each final model is evaluated on the previously untouched testing set. We visualize model performance across the distribution of testing data with a scatter diagram and quantile-quantile plot for which successful model predictions should lie close to a 1:1 line.

These trained models can then easily be deployed on the on-board processing computer so that during subsequent surveys, target concentrations can be inferred as imagery are collected and processed. The application of each model to the collected hyperspectral data cubes results in a map of the distribution of each water quality variable across the pond.

@@

\begin{figure}
\includegraphics[width=0.7\columnwidth]{paper/figures/results/downwelling-hist.pdf}
\caption{Distribution of total downwelling intensity during each of the three HSI collection flights. The multi-modal nature of these distributions reflects the impact of the relative orientation of the drone to the sun as well as potential occlusion due to the presence of clouds.\label{fig:downwelling-hist}}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{paper/figures/results/fits/physical-fitres.png}
\caption{Scatter diagrams (left) and quantile-quantile plots (right) for the hyperparameter optimized RFR models for the physical variables measured by the USV.\label{fig:physical-fit}}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{paper/figures/results/fits/physical-ranking.pdf}
\caption{Ranked permutation importance for each feature in the physical variable models. Permutation importance measured the decrease in model $R^2$ value after replacing each feature in the prediction set with a random permutation of its values.\label{fig:physical-fi}}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{paper/figures/results/maps/physical.pdf}
\caption{Maps generated by applying each of the physical variable models to the hyperspectral data cubes collected on 11-23. Overlaid over the predictions are color-filled squares showing the associated in-situ reference data for the same collection period. The size of the squares has been exaggerated for the visualization. We note that there is good agreement between the model predictions and the reference data. \label{fig:map-physical}}
\end{figure}

@@

>>>>

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{
Inland waters pose a unique challenge for water quality monitoring by remote sensing techniques due to their complicated spectral features and small-scale variability. At the same time, collecting the high-quality reference data needed to calibrate remote sensing data products is both time consuming and expensive. In this study, we present the further development of a robotic team composed of an uncrewed surface vessel (USV) providing in situ reference measurements and an unmanned aerial vehicle (UAV) equipped with a hyperspectral imager. Together, this team is able to address the limitations of existing approaches by enabling the simultaneous collection of comprehensive in situ data together with high-resolution hyperspectral imagery. We showcase the capabilities of this team using data collected in a north Texas pond during three collections in 2020. Machine learning models for 13 variables are trained using the data set of paired in situ measurements and coincident reflectance spectra. These models can successfully estimate physical variables, including temperature, conductivity, pH, and turbidity, as well as the concentrations of blue-green algae, colored dissolved organic matter (CDOM), chlorophyll-a, crude oil, optical brighteners and the ions $\mathrm{Ca}^{2+}$, $\mathrm{Cl}^{-}$, and $\mathrm{Na}^{+}$. We extend the training procedure to utilize conformal prediction to estimate 90\% confidence intervals for the output of each trained model. Maps generated by applying each trained model to the collected images reveal the small-scale spatial variability within the pond. Additionally, the permutation importance computed for each feature of the trained models highlights the spectral features relevant to each water quality variable.
}

@@

For decades, remote sensing imagery has been used for environmental monitoring with applications ranging from resource mapping, land type classification, and urban growth assessment to wildfire tracking, natural disaster tracking, and many more \cite{melesse2007remote, joyce2009review}. Among these applications, the retrieval of water quality variables from remote sensing imagery remains challenging due to the difficulty of obtaining in situ reference data that coincide with available satellite imagery. Many studies have shown success modeling relevant water quality variables such as colored dissolved organic matter (CDOM), cholorophyll-a, and total suspended sediment concentrations by using combinations of spectral bands from terrestrial remote sensing satellites. However, the in situ reference data used in these studies typically require the time-consuming collection of individual samples for lab analysis or continuous monitoring at fixed sites \cite{brezonik2005landsat, paavel2011optical, gursoy2019investigating, bonansea2019using, sagan2020monitoring, remote-sensing-finland}.

Much effort has recently been spent on the curation of comprehensive datasets combining water quality records with decades of satellite imagery to enable the development of new methods for monitoring water quality. For example, Aurin et al. curated over 30 years of oceanographic field campaign data with associated coincident satellite imagery \cite{aurin2018remote}. Similarly, Ross et al. have combined more than 600,000 records of dissolved organic carbon, chlorophyll-a and other variables of water quality together with Landsat historical reflectance data for the period 1984--2019 \cite{ross2019aquasat}. With a sufficient quantity of data, machine learning methods can augment traditional approaches using reflectance band ratios to enable estimation of water quality variables and perform water quality classification \cite{schiller1999neural, brown2008neural, lary2010artificial, feng2015urban,belgiu2016random}. However, the low spatial and spectral resolution of available multiband remote sensing satellites makes it difficult to analyze inland waters with small spatial features and complicated spectral signatures. Furthermore, the long return times of these satellites limit their use for real-time monitoring, for example, during oil spills \cite{fingas2017review}.

@@

Despite the increased spectral and spatial resolution afforded by UAV platforms, these improvements alone cannot address the limited spatial resolution of in situ reference data.  To address this gap and enable comprehensive real-time water quality \textcolor{red}{detection}, we have developed a robotic team composed of an autonomous uncrewed surface vessel (USV) equipped with a collection of reference grade instruments together with a UAV carrying a hyperspectral imager \cite{robotTeam1}. By incorporating these reference instruments on a maneuverable USV platform, we are able to rapidly collect large volumes of water quality data for a variety of physical, biochemical, and chemical variables. Critically, the USV enables the collection of reference data with significantly improved spatial resolution compared to other approaches. These in situ measurements then provide the ground-truth data used to train machine learning models which map reflectance spectra captured by the hyperspectral imager to the desired water quality variables.

In this study, we expand on our previous work by demonstrating the capabilities of this coordinated robot team using data collected at a pond in North Texas over three separate days in 2020. Specifically, we show the ability to accurately predict physical variables such as temperature, conductivity, pH, and turbidity, as well as biochemical constituents, including blue-green algae pigments, CDOM, and chlorophyll-a, in addition to concentrations of crude oils, optical brighteners, and a variety of ions. In addition, the considerable size of the collected data set allows us to quantify the uncertainty in the resulting machine learning models by using conformal prediction techniques. This methodology addresses the shortcomings of existing approaches by facilitating rapid, real-time data collection coupled with on-board hyperspectral image processing and machine learning model deployment with uncertainty quantification. The resulting models are small enough to be deployed on board the UAV to enable real-time streaming of data products to the ground. This approach drastically reduces the time from data collection to insight.

\textcolor{red}{The primary goal of this study is to demonstrate the capabilities of a coordinated robotic team combining in situ sensing together with hyperspectral imaging for the rapid characterization of water composition. Specifically, we seek to address the primary challenge of acquiring remote sensing data products coincident with reference measurements that are critical for the retrieval of water composition and water quality parameters. By bringing together
- there are three key goals in this study: The first is to demonstrate the variety of water quality and composition variables that can be predicted from our collected hypserspectral imager. The second is to demonstrate that hyperspectral imagery from multiple observations can by combined despite variability in lighting conditions, and the third is to introduce a method for uncertainty quantification which is crucial for the evaluation of model predictions. 
}

@@

\begin{figure}[t!]
\includegraphics[width=\columnwidth]{paper/figures/materials-and-methods/annotated-boat.pdf}
\caption{ Configuration of the USV: (\textbf{a}) Frontal view of the USV showing the Eureka Manta+40 Multiprobes mounted on the underside of the boat. (\textbf{b}) The USV deployed in the water.\label{fig:boat-components}}
\end{figure}

@@

\begin{table}[t!] 
\caption{In-situ reference sensors utilized in this study.\label{tab:sensors}}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{CCCCC}
\toprule
\textbf{Sensor}	& \textbf{Units} & Resolution & \textbf{Sensor Type} & \textbf{Target Category}\\
\midrule
Temperature		                    & $^{\circ}$C        & 0.01  & Thermistor                           & Physical \\
Conductivity                        & $\mu$S/cm          & 0.01  & Four-Electrode Graphite Sensor       & Physical  \\
pH                                  & logarithmic (0-14) & 0.01  & Flowing-junction Reference Electrode & Physical \\
Turbidity                           & FNU                & 0.01  & Ion-Selective Electrode              & Physical \\
$\mathrm{Ca^{2+}}$                  & mg/l               & 0.1   & Ion-Selective Electrode              & Ions \\
$\mathrm{Cl^-}$                     & mg/l               & 0.1   & Ion-Selective Electrode              & Ions \\
$\mathrm{Na^+}$                     & mg/l               & 0.1   & Ion-Selective Electrode              & Ions \\
Blue-Green Algae (phycoerythrin)    & ppb                & 0.01  & Fluorometer                          & Biochemical \\
Blue-Green Algae (phycocyanin)      & ppb                & 0.01  & Fluorometer                          & Biochemical \\
CDOM                                & ppb                & 0.01  & Fluorometer                          & Biochemical \\
Chlorophyll A                       & ppb                & 0.01  & Fluorometer                          & Biochemical \\
Optical Brighteners                 & ppb                & 0.01  & Fluorometer                          & Chemical \\
Crude Oil                           & ppb                & 0.01  & Fluorometer                          & Chemical\\
\bottomrule
\end{tabularx}
% \noindent{\footnotesize{\textsuperscript{1} Tables may have a footer.}}
\end{table}

As shown in Table~\ref{tab:sensors}, the physical measurements and ion sensors are largely based on different electrode configurations, while the chemical and biochemical measurements are all optically significant in UV and visible light, enabling their determination by fluorometry \cite{de2007ion,trees2002fluorometric,tillman2017evaluation}. The pigments phycoerythrin and phycocyanin are used to determine the blue-green algae content, which together with chlorophyll-a enable us to assess the distribution of photosynthetic life in the pond \cite{Brient2008APP, boyer2009phytoplankton}. In addition, we also measured the concentration of colored dissolved organic matter (CDOM), which impacts light penetration and serves as a primary source of bioavailable carbon. Crude oil (natural petroleum) and optical brightener concentrations are also measured with fluorometers and are relevant for identifying sources of industrial contamination, natural seepage, and sewage \cite{brown2003review,cao2009evaluation}.

\textcolor{red}{Should we add something here about non-optically active substances?}

@@

\begin{figure}[b!]
\includegraphics[width=\columnwidth]{paper/figures/materials-and-methods/annotated-drone.pdf}
\vspace{-0.2in}
\caption{Configuration of the UAV: (\textbf{a}) the hyperspectral imager and acquisition computer. (\textbf{b}) the assembled UAV with secondary processing computer and (upward facing) downwelling irradiance spectrometer. \label{fig:drone-components}}
\end{figure}

@@

\begin{figure}[t!]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\vspace{-0.2in}
\includegraphics[width=14cm]{paper/figures/materials-and-methods/pipeline-figure-2.pdf}
\end{adjustwidth}
\vspace{-0.2in}
\caption{Hyperspectral Image Processing: Hyperspectral data cubes are collected one scan-line at a time (left). By utilizing downwelling irradiance spectra we convert each pixel from spectral radiance to reflectance. By using orientation and position data from the on-board GPS and INS, we georeference each pixel to assign it a latitude and longitude on the ground. The final data product is a georectified hyperspectral reflectance data cube visualized at the right. \label{fig:hsi-pipeline}}
\end{figure}

@@

\begin{figure}[t!]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\vspace{-0.2in}
\includegraphics[width=15.5cm]{paper/figures/materials-and-methods/HyperSpectralInfoGraphic.pdf}
\end{adjustwidth}
\caption{A georectified reflectance data cube is visualized (center) with the $\log10$-reflectance along the z axis and a pseudo-color image on the top. In the top left we visualize the downwelling irradiance spectrum (the incident light). The surrounding plots showcase exemplar pixel reflectance spectra for open water, dry grass, algae, and a rhodamine dye plume used to test the system.\label{fig:hsi-infographic}}
\end{figure}

@@

Data were collected using the robot team at a site in North Texas during three observations on 11-23-2020, 12-09-2020, and 12-10-2020. For each collection period, the UAV first completed a broad survey of the pond, capturing multiple hyperspectral data cubes. Subsequently, the USV sampled across the pond, collecting in situ reference measurements. Each of these reference measurements was then collocated with individual pixel spectra, where the USV track overlapped with UAV pixels. To account for any time lag in the values measured by the in situ instruments and to account for the USV size in comparison to the data cube's spatial resolution, each in-situ measurement is associated with a 3$\times$3 grid of HSI pixels, that is, a 30 cm $\times$ 30 cm square. These combined data form the tabular data set on which we train regression models; pixel spectra form input features, and each separate USV sensor forms a distinct target variable.

Each collection was performed near solar noon to maximize the amount of incident sunlight illuminating the water. For the site in North Texas, this corresponded to average solar zenith angles of 54.9$^{\circ}$, 56.7$^{\circ}$, and 56.75$^{\circ}$ for 11-23-202, 12-09-2020, and 12-10-2020 respectively. Given the hyperspectral imager acquires data cubes at nadir, there was little concern for effects due to sunglint. However, to account for any potential variation in lighting conditions between data cubes, we augment the training set with additional features including the drone viewing geometry (roll, pitch, heading) and solar illumination geometry (solar azimuth, solar elevation, solar zenith) as well as the total at-pixel intensity before reflectance conversion, the total downwelling intensity, and the drone altitude. Further feature engineering is performed to add additional spectral indices that utilize combinations of specific wavelength bands such as the normalized difference vegetation index (NDVI), normalized difference water index (NDWI), simple ratio (SR), photochemical reflectance index (PRI) and more, as outlined in \cite{envi_vegetation_indices, thenkabail2018hyperspectral,kaufman1992atmospherically, SpectralIndexWheat}. A comprehensive list of these added features is provided in Supplementary Table S1. The final dataset includes a total of 526 features (462 reflectance bands plus 64 additional features) with over 120,000 records.

@@

For each of the 13 target variables listed in Table~\ref{tab:sensors}, the data was randomly partitioned into a 90:10 training/testing split. To model the data, we chose to use the Random Forest Regressor (RFR) as implemented in the Machine Learning Framework for Julia package,  MLJ \cite{MLJ1}.  Random forests are an ensembling technique based on bagged predictions of individual decision tree regressors trained using the classification and regression trees (CART) algorithm. Each tree in a RFR is trained on a random subset of features \textit{and} a random subset of training records \cite{decision-trees, random-forest}. Random Forests are particularly attractive due to their fast training and inference times. Furthermore, studies continue to observe that tree-based models like the Random Forest remain superior for tabular datasets \cite{grinsztajn2022tree, shwartz2022tabular}.

As reflectance values in adjacent wavelength bins tend to correlate with each other and, therefore, may not necessarily contribute additional information content to the final model, it is desirable to evaluate the relative importance of each feature to the trained model predictions. This is useful both for identifying the most relevant features and for performing feature selection to reduce the final model size. By default, tree-based methods, such as the RFR allow for an impurity-based ranking as described in \cite{random-forest, rfr-importance-ranking}. However, these methods have been shown to be biased towards high cardinality and correlated features \cite{strobl2008conditional}. Therefore, we choose to use the model-agnostic permutation importance as described by \cite{parr2018beware}. To do this, we further partition the training data set, resulting in an 80:10:10 split with 80\% of points used for model training, 10\% of points used for validation and determination of feature importance, and the final 10\% held out as an independent testing set. The importance of the $j^{\text{th}}$ feature is then computed as
\begin{equation}
    \text{Imp}_j = R^2(f(X_{\text{val}}), y_{\text{val}}) - R^2(f(X_{\text{val}}^{(j)}), y_{\text{val}})
\end{equation}
where $(X_{\text{val}}, y_{\text{val}})$ is the validation set, $f(\cdot)$ is the trained model, $R^2(\cdot, \cdot)$ is the coefficient of determination, and $X_{\text{val}}^{(j)}$ is validation feature set with the $j^{th}$ column randomly permuted. The permutation importance is therefore understood as the decrease in model performance when the $j^{th}$ feature is replaced by random values from the validation set.

To assess the uncertainty of the final model's predictions, we employed Inductive Conformal Prediction as described in \cite{conformal-prediction-1, conformal-prediction-2, conformal-prediction-3, conformal-prediction-4}. To do this, we computed a set of nonconformity scores of the trained model on the validation set using an uncertainty heuristic, in this case, the absolute error
\begin{equation}
    s_i = s(X_i, y_i) = \lvert f(X_i) - y_i \rvert = \lvert \hat{y}_i - y_i \rvert
\end{equation}
where $f(X_i)=\hat{y}_i$ denotes the trained ML model applied to the $i^{th}$ calibration record. These n-many scores are sorted, and the interval half width, $d$, is calculated as the $\frac{\lceil(1-\alpha)(n+1) \rceil}{n}$ quantile of this set in order to achieve a coverage of $1-\alpha$ on the calibration set. Prediction intervals for new data are then formed as $f(X)\pm d$. For this study, we chose $\alpha=0.1$ for a coverage corresponding to a 90\% confidence interval.

Using these tools, the training procedure for each model was as follows: First, each model was trained using six-fold cross-validation on the full 526-feature training set with default hyperparameter values. Feature importances for the trained model were then computed, and the top 25 features were identified. A second model was then trained using the same 6-fold cross-validation scheme with \textit{only} these 25 most important features together with hyperparameter optimization using a random search over the number of trees and sampling fraction. The number of trees was optimized as it has a significant impact on both model performance and inference time. The sampling ratio determines the fraction of training records that each individual tree is exposed to during training. Tuning this parameter helps limit overfitting by increasing the diversity trees in the ensemble. Additionally, we chose to fix the maximum tree depth to 20 to control the final model size such that each trained model can fit in-memory on the on board UAV processing computer. The remaining hyperparameters were left to their default values in order to constrain the total optimization space. The performance of each model was evaluated by computing out-of-fold scores for the coefficient of determination, as well as the
\begin{align}
    \text{RMSE} &= \sqrt{\frac{\sum\limits_{i=1}^N (\hat{y}_i-y_i)^2}{N}}, \\
    \text{MAE} &= \frac{\sum\limits_{i=1}^N \lvert \hat{y}_i - y_i \vert}{N},
\end{align}
where RMSE is the root-mean-square error, and MAE is the mean absolute error between true measurements $y_i$ and predictions $\hat{y}_i$.

Having identified the model with the best out-of-fold performance, we proceeded to train the final hyperparameter optimized model on the full training set with associated uncertainty estimated using conformal prediction.  Then each final model was evaluated on the previously untouched testing set. We visualize model performance across the distribution of testing data with a scatter diagram and quantile-quantile plot for which successful model predictions should lie close to a 1:1 line.

These trained models can then easily be deployed on the on-board processing computer so that during subsequent surveys, target concentrations can be inferred as imagery are collected and processed. The application of each model to the collected hyperspectral data cubes results in a map of the distribution of the water composition across the pond.

@@

\begin{figure}[t!]
\vspace{-0.2in}
\includegraphics[width=0.7\columnwidth]{paper/figures/results/downwelling-hist.pdf}
\vspace{-0.2in}
\caption{Distribution of total downwelling intensity during each of the three HSI collection flights. The multi-modal nature of these distributions reflects the impact of the relative orientation of the drone to the sun as well as potential occlusion due to the presence of clouds.\label{fig:downwelling-hist}}
\end{figure}

@@

\begin{figure}[p]
\centering
\includegraphics[width=\columnwidth]{paper/figures/results/fits/physical-fitres.png}
\caption{Scatter diagrams (left) and quantile-quantile plots (right) for the hyperparameter optimized RFR models for the physical variables measured by the USV.\label{fig:physical-fit}}
\end{figure}

@@

\begin{figure}[t!]
\centering
\vspace{-0.2in}
\includegraphics[width=\columnwidth]{paper/figures/results/fits/physical-ranking.pdf}
\caption{Ranked permutation importance for each feature in the physical variable models. Permutation importance measured the decrease in model $R^2$ value after replacing each feature in the prediction set with a random permutation of its values.\label{fig:physical-fi}}
\end{figure}

@@

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{paper/figures/results/maps/physical.pdf}
\caption{Maps generated by applying each of the physical variable models to the hyperspectral data cubes collected on 11-23. Overlaid over the predictions are color-filled squares showing the associated in-situ reference data for the same collection period. The size of the squares has been exaggerated for the visualization. We note that there is good agreement between the model predictions and the reference data. \label{fig:map-physical}}
\end{figure}

@@

% \cite{cohen2004landsat, melesse2007remote, allison2016airborne, joyce2009review, robert2017analysis}
